---
title: "STRUCTURE-like Ancestry Analysis with sNMF & DAPC (from genind SNPs) - MS365 Copilot GPT5"
author: "`r Sys.info()[['user']]`"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  pdf_document:
    toc: true
    number_sections: true
params:
  genind_rds: "genind.rds"
  missing_ind: 0.20
  missing_loc: 0.10
  maf_min: 0.01
  K_min: 1
  K_max: 12
  repetitions: 10
  alpha: 100
  seed: 2025
  save_prefix: "sNMF"
  run_dapc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = FALSE, fig.width = 9, fig.height = 5)

sc.base <- "W:/9_PBT/01_chum/reference_databases/bcm_SNP_coastwide_v.5.0.0_2022-08-18/cmStockCodesCU.txt"
repunits.base <- "W:/9_PBT/01_chum/reference_databases/bcm_SNP_coastwide_v.5.0.0_2022-08-18/Haida_Gwaii/repunits_full.txt"




need <- c("adegenet", "LEA", "dplyr", "ggplot2", "tidyr", "readr", "stringr", "scales")
to_install <- need[!sapply(need, requireNamespace, quietly = TRUE)]
if (length(to_install)) install.packages(to_install)
lapply(need, library, character.only = TRUE)

set.seed(params$seed)
dir.create("outputs", showWarnings = FALSE)
dir.create(file.path("outputs","allK_sNMF"), showWarnings = FALSE, recursive = TRUE)
dir.create(file.path("outputs","allK_DAPC"), showWarnings = FALSE, recursive = TRUE)
cat("Outputs will be saved to: outputs/\n")

# Helper: STRUCTURE-like barplot with population labels, full boundary lines, and repunit labels on top
plot_structure_like <- function(Q_ord, pop_ord, joined, title, ylab, outfile, colpal,
                                label_cex = 0.7, repunit_line_lwd = 2,
                                tick_frac = 0.06, top_tick_on_repunit = TRUE,
                                bottom_ticks = TRUE) {
  stopifnot(all(c("collection","repunit") %in% names(joined)))

  png(outfile, width = 2400, height = 850, res = 160)
  op <- par(mar = c(10, 4, 7, 1), xpd = NA)

  bp <- barplot(t(Q_ord), col = colpal, space = 0, border = NA,
                xlab = "", ylab = ylab, main = title, xaxt = "n")

  # Population boundaries and labels
  counts     <- as.integer(table(pop_ord))
  ends_idx   <- cumsum(counts)
  starts_idx <- c(1, head(ends_idx, -1) + 1)
  centers_vec <- tapply(bp, pop_ord, mean)
  pop_labels  <- names(centers_vec)

  # Axis labels (no baseline/ticks)
  axis(1, at = as.numeric(centers_vec), labels = pop_labels,
       las = 2, cex.axis = label_cex, tick = FALSE, lwd = 0, lwd.ticks = 0)

  # All population boundaries
  sep_x <- if (length(ends_idx) > 1)
             (bp[ends_idx[-length(ends_idx)]] + bp[starts_idx[-1]]) / 2
           else numeric(0)

  # Repunit membership for each population label, then find change points
  coll_to_rep <- setNames(joined$repunit, joined$collection)
  pop_rep <- unname(coll_to_rep[as.character(pop_labels)])
  pop_rep[is.na(pop_rep)] <- "Unassigned"
  rep_changes <- which(pop_rep[-length(pop_rep)] != pop_rep[-1])
  sep_rep_x <- if (length(rep_changes)) sep_x[rep_changes] else numeric(0)

  # Tick geometry
  usr <- par("usr"); y0 <- usr[3]; y1 <- usr[4]
  h  <- tick_frac * (y1 - y0)   # height
  pad <- 0.01 * (y1 - y0)

  # Bottom ticks at ALL population boundaries (optional)
  if (bottom_ticks && length(sep_x)) {
    segments(x0 = sep_x, y0 = y0, x1 = sep_x, y1 = y0 + h,
             col = "grey40", lwd = 1.2)
  }

  # Repunit bands + labels (top)
  pop_left  <- bp[starts_idx] - 0.5
  pop_right <- bp[ends_idx] + 0.5
  r <- rle(pop_rep)
  rep_blocks <- data.frame(repunit = r$values, len = r$lengths)
  rep_blocks$pop_start_index <- c(1, head(cumsum(rep_blocks$len), -1) + 1)
  rep_blocks$pop_end_index   <- cumsum(rep_blocks$len)
  rep_blocks$x_left  <- pop_left[rep_blocks$pop_start_index]
  rep_blocks$x_right <- pop_right[rep_blocks$pop_end_index]
  rep_blocks$x_mid   <- (rep_blocks$x_left + rep_blocks$x_right)/2

  y_line <- y1 + 0.02 * diff(usr[3:4])   # band line a bit above panel
  y_text <- y1 + 0.12 * diff(usr[3:4])   # label above the band

  segments(rep_blocks$x_left, y_line, rep_blocks$x_right, y_line,
           xpd = NA, lwd = repunit_line_lwd, col = "black")
  text(rep_blocks$x_mid, y_text, labels = rep_blocks$repunit,
       xpd = NA, cex = 0.9)

  # TOP ticks only at repunit boundaries
  if (top_tick_on_repunit && length(sep_rep_x)) {
    # Option A: up above panel
    segments(x0 = sep_rep_x, y0 = y1, x1 = sep_rep_x, y1 = y1 + h,
             col = "grey20", lwd = 1.4)

    # Option B (comment above and uncomment below) to stop just below the band:
    # segments(x0 = sep_rep_x, y0 = y1, x1 = sep_rep_x, y1 = y_line - pad,
    #          col = "grey20", lwd = 1.4)
  }

  par(op); dev.off()
}

library(adegenet)
library(LEA)
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(stringr)
library(scales)






load_genepop <- function(datatype = "SNP"){
  
  # Manually choose files (function depends windows or linux)
  print("Select your genepop from file...")
  if(.Platform$OS.type == "unix") {
    
    my_genepop.path <- tk_choose.files(caption = "Select a genepop file" )
    
  } else if(.Platform$OS.type == "windows") {
    
    my_genepop.path <- choose.files( default=file.path("02_raw_data/")
                                     , caption = "Select a genepop file")
    
  }
  
  
  # Set allele.code
  if(datatype=="SNP"){
    allele.code <- 2
  }else if(datatype=="microsat"){
    allele.code <- 3
  }
  
  # Read in data
  print(paste0("Loading genepop from ", my_genepop.path))
  
  obj <- read.genepop(file = my_genepop.path
                      , ncode = allele.code)
  
  print(obj)
  
  assign(x = "obj", value = obj, envir = .GlobalEnv)
  assign(x = "datatype", value = datatype, envir = .GlobalEnv)
  assign(x = "my_genepop.path", value = my_genepop.path, envir = .GlobalEnv)
  
}



update_pop_names <- function(df = obj, sep_by = "collection", name_by = "stockname", add_CU = FALSE){
  
  # Create backup
  obj.bck <- df
  
  # Extract genepop indiv names and put into df
  indiv_names.df <- as.data.frame(rownames(df$tab))
  colnames(indiv_names.df) <- "indiv.name"
  head(indiv_names.df)
  
  # Separate the df into individual components
  indiv_names.df <- separate(data = indiv_names.df
                             , sep = "_"
                             , col = "indiv.name"
                             , into = c("stock.code", "year", "fish.id", "sex")
                            )
  head(indiv_names.df)
  dim(indiv_names.df)

  
  ## Generate new names for populations
  # Reporting
  print(paste0("Renaming pops to ", sep_by))
  
  if(sep_by=="collection"){
    
    pop_short <- str_pad(string = indiv_names.df$stock.code, width = 2, side = "left", pad = "0")
    
  }else if(sep_by=="collection_and_year"){
    
    pop_short <- paste0(str_pad(string = indiv_names.df$stock.code, width = 2, side = "left", pad = "0"), "_", indiv_names.df$year)
    
  }else {
    
    print("Not renaming pops")
    pop_short <- pop(df)
    
  }

  ## Update pop names
  pop(df) <- pop_short
  print(unique(pop(df)))
  
  ## Collect the new names into a dataframe
  pop_short.df <- as.data.frame(pop(df))
  colnames(pop_short.df) <- "pop_short"
  
  # Make numeric for matching with stock code below
  if(sep_by == "collection"){
    
    # TO REMOVE
    #pop_short.df$pop_short <- as.integer(as.character(pop_short.df$pop_short)) # for below, this must be integer
    # TO REMOVE
    
  }else if(sep_by == "collection_and_year"){
    
    # separate the column into the collection and the year again
    pop_short.df <- separate(data = pop_short.df, sep = "_", col = "pop_short", into = c("stock.code", "year"))
    
    # rename stock.code into 'pop_short' for compatibility between methods
    colnames(pop_short.df)[which(colnames(pop_short.df)=="stock.code")] <- "pop_short"
    
  }else{
    
    print("Pop names still in original format")
    
  }
  
  # To merge with stock code file below, the pop short variable must be an integer
  if(sep_by != "none"){
    pop_short.df$pop_short <- as.integer(as.character(pop_short.df$pop_short)) 
  }

  
  ## Change from numeric to names
  #### 01.2 Pop names as numeric or character?
  if(sep_by!="none" && name_by=="stockname"){
    
    # Generate an order column for re-ordering the samples' stock code vector
    pop_short.df$id <- 1:nrow(pop_short.df)
    
    # Reporting
    print("Reading in stock codes")
    
    stock_codes.df <- read.delim2(file = sc.base, stringsAsFactors = F)
    
    # Safety check - are there duplicate values?
    print("The following stock codes are seen more than once:   ")
    colnames(stock_codes.df)[which(colnames(stock_codes.df)=="StockCode")] <- "Code" #Temp solution to new header names
    stock_codes.df[duplicated(x = stock_codes.df$Code), "Code"]
    print("If there are duplicated values, this will cause the program to crash")
    
    # Merge to get the stock code name
    rosetta <- merge(x = stock_codes.df, y = pop_short.df, by.x = "Code", by.y = "pop_short", sort = F, all.y = T)
    rosetta$collection <- as.character(rosetta$collection)
    rosetta$repunit <- as.character(rosetta$repunit)
    rosetta$ProvState <- as.character(rosetta$ProvState)
    str(rosetta)
    
    print("re-ordering back into original order")
    rosetta <- rosetta[order(rosetta$id), ]
    
    # Add the year if necessary, and create vector for using as pop ID
    if(sep_by=="collection_and_year"){
      
      rosetta$name <- paste0(rosetta$collection, "-", rosetta$year)
      
    }else if(sep_by=="collection"){
      
      rosetta$name <- rosetta$collection
      
    }
    
    # Add the CU if requested
    if(add_CU==TRUE){
      
      print("Adding CU to the collection name for the pop variable")
      
      rosetta$name <- paste(rosetta$name, rosetta$repunit, sep =  "_")
      
    }
    
    # Use the newly designated identifier per individual
    pop(df) <- rosetta$name
    
    # Reporting
    print(unique(pop(df)))
    print(table(pop(df)))
    
  } else {
    
    print("Not renaming pop stock code to stock names")
    
  }
  
  assign(x = "obj", value = df, envir = .GlobalEnv)
  assign(x = "obj.bck", value = obj.bck, envir = .GlobalEnv)
  assign(x = "sep_by", value = sep_by, envir = .GlobalEnv)
  assign(x = "name_by", value = name_by, envir = .GlobalEnv)
  assign(x = "add_CU", value = add_CU, envir = .GlobalEnv)
  
}


load_genepop()

update_pop_names(sep_by = "collection", name_by = "stockname",add_CU = FALSE)



g <- obj



repunits <- read.delim(file=repunits.base)
stockcodes<- read.delim(file=sc.base)


joined <- merge(repunits,stockcodes,by="repunit",all.y=TRUE)



```

# 1. Load `genind` and Basic Checks

```{r load-genind}
if (!exists("g")) {
  stopifnot(file.exists(params$genind_rds))
  g <- readRDS(params$genind_rds)
}
stopifnot(inherits(g, "genind"))
cat("Genotype type:", g@type, "\n")
cat("Ploidy:", ploidy(g), "\n")
cat("Individuals:", nInd(g), "| Loci:", nLoc(g), "\n")
if (is.null(pop(g))) { pop(g) <- factor(rep("All", nInd(g))) }
```

---

# 2. QC (Missingness & MAF) and Build Dosage Matrix

```{r qc}
# Build a SNP dosage matrix (0/1/2) directly from genind
M <- tab(g, freq = FALSE, NA.method = "asis")  # nInd x (2 * nLoc)
cn <- colnames(M)
split_last_dot <- function(x) { i <- max(gregexpr("\\.", x)[[1]]); c(substr(x,1,i-1), substr(x,i+1,nchar(x))) }
sp <- do.call(rbind, lapply(cn, split_last_dot))
loci <- sp[,1]; alle <- sp[,2]
loc_levels <- unique(loci)
idx_alt <- unlist(lapply(loc_levels, function(L) which(loci==L & alle==sort(alle[loci==L])[2])[1]))
dosage <- M[, idx_alt, drop = FALSE]
# genlight
gl <- new("genlight", as.matrix(dosage))
ploidy(gl) <- ploidy(g); indNames(gl) <- indNames(g); locNames(gl) <- loc_levels; pop(gl) <- pop(g)

# Missingness filters on gl
Xi <- as.matrix(gl)
ind_miss <- rowMeans(is.na(Xi)); keep_ind <- ind_miss <= params$missing_ind; gl <- gl[keep_ind, ]
X <- as.matrix(gl)
loc_miss <- colMeans(is.na(X)); keep_loc <- loc_miss <= params$missing_loc; gl <- gl[, keep_loc]
X <- as.matrix(gl)
p <- colMeans(X, na.rm = TRUE)/2; maf <- pmin(p, 1-p); keep_maf <- maf >= params$maf_min; gl <- gl[, keep_maf]
X <- as.matrix(gl)
cat(sprintf("After QC: %d individuals, %d SNPs retained.\n", nInd(gl), nLoc(gl)))

# ---- Create DAPC input directly from filtered gl (avoid subsetting g by names) ----
# Mean-impute per locus, center columns, drop zero-variance columns
X_imp <- X
cm <- colMeans(X_imp, na.rm = TRUE)
for (j in seq_len(ncol(X_imp))) {
  nas <- is.na(X_imp[, j]); if (any(nas)) X_imp[nas, j] <- cm[j]
}
X_center <- sweep(X_imp, 2, colMeans(X_imp), FUN = "-")
keep_var <- apply(X_center, 2, function(z) sd(z) > 0)
X_qc <- X_center[, keep_var, drop = FALSE]
# Set row/col names for traceability
rownames(X_qc) <- indNames(gl); colnames(X_qc) <- locNames(gl)[keep_var]
stopifnot(all(is.finite(X_qc)))
```

---

# 3. sNMF across K, select bestK, and store **all K** results

```{r snmf-run, cache=TRUE}
library(LEA)
X_lea <- X; X_lea[is.na(X_lea)] <- 9; storage.mode(X_lea) <- "integer"
geno_file <- file.path("outputs", paste0(params$save_prefix, "_data.geno"))
LEA::write.geno(X_lea, geno_file)
K_range <- params$K_min:params$K_max
project <- LEA::snmf(geno_file, K = K_range, repetitions = params$repetitions, entropy = TRUE, alpha = params$alpha, project = "new")
ce_df <- lapply(K_range, function(k){ runs <- seq_len(params$repetitions); ce_runs <- sapply(runs, function(r) LEA::cross.entropy(project, K=k, run=r)); data.frame(K=k, run=runs, cross_entropy=ce_runs) }) %>% dplyr::bind_rows()
ce_best <- ce_df %>% dplyr::group_by(K) %>% dplyr::summarise(cross_entropy = min(cross_entropy), .groups = "drop")
bestK <- ce_best$K[which.min(ce_best$cross_entropy)]
# store best run per K and Q matrices
best_run_perK <- sapply(K_range, function(k) which.min(sapply(seq_len(params$repetitions), function(r) LEA::cross.entropy(project, K=k, run=r))))
Q_list <- lapply(K_range, function(k){ Qk <- LEA::Q(project, K=k, run=best_run_perK[match(k, K_range)]); rownames(Qk) <- indNames(gl); colnames(Qk) <- paste0("K", seq_len(k)); Qk })
names(Q_list) <- as.character(K_range)
# plot CE
ce_plot <- ggplot(ce_best, aes(x=K,y=cross_entropy)) + geom_line(color="#2c7fb8") + geom_point(size=2,color="#2c7fb8") + geom_vline(xintercept=bestK, linetype=2, color="firebrick") + labs(title="sNMF Model Selection", y="Cross-entropy (best run per K)") + theme_bw(base_size=12)
ggsave(file.path("outputs", sprintf("%s_sNMF_cross_entropy.png", params$save_prefix)), ce_plot, width=7.5, height=4, dpi=150)
readr::write_csv(ce_df, file.path("outputs", sprintf("%s_sNMF_cross_entropy_all_runs.csv", params$save_prefix)))
```

---

# 4. Export Q for bestK, align ordering by `joined`, and plot (main + **all K** to subfolder)

```{r qmatrix-and-plots}
stopifnot(exists("joined"))
stopifnot(all(c("Display_Order","collection","repunit") %in% names(joined)))
joined <- joined[order(joined$Display_Order, joined$collection), ]
ordered_pops <- unique(joined$collection)
popvec <- as.character(pop(gl))
missing_in_joined <- setdiff(unique(popvec), ordered_pops)
if (length(missing_in_joined)) { warning(sprintf("These populations are in popvec but not in `joined$collection`: %s", paste(missing_in_joined, collapse=", "))); pop_levels <- c(ordered_pops, setdiff(unique(popvec), ordered_pops)) } else { pop_levels <- ordered_pops }
popfac <- factor(popvec, levels = pop_levels)
ord <- order(popfac); pop_ord <- popfac[ord]
colpal_byK <- lapply(K_range, function(k) if (requireNamespace("scales", quietly=TRUE)) scales::hue_pal()(k) else rainbow(k)); names(colpal_byK) <- as.character(K_range)
Q <- Q_list[[as.character(bestK)]]
readr::write_csv(as.data.frame(Q) %>% tibble::rownames_to_column("Individual"), file.path("outputs", sprintf("%s_Qmatrix_K%d.csv", params$save_prefix, bestK)))
assign_k <- max.col(Q, ties.method="first"); max_q <- apply(Q,1,max)
assignments <- data.frame(Individual=rownames(Q), Population=as.character(popfac), Cluster=paste0("K",assign_k), MaxQ=max_q)
readr::write_csv(assignments, file.path("outputs", sprintf("%s_assignments_K%d.csv", params$save_prefix, bestK)))
# plot bestK
Q_best_ord <- Q[ord,,drop=FALSE]; colpal <- colpal_byK[[as.character(bestK)]]
plot_structure_like(Q_best_ord, pop_ord, joined, sprintf("sNMF Q-matrix (K = %d)", bestK), "Ancestry coefficient", file.path("outputs", sprintf("%s_barplot_grouped_K%d.png", params$save_prefix, bestK)), colpal)
# plot all K
for (k in K_range) { Qk <- Q_list[[as.character(k)]]; if (is.null(Qk)) next; Qk_ord <- Qk[ord,,drop=FALSE]; plot_structure_like(Qk_ord, pop_ord, joined, sprintf("sNMF Q-matrix (K = %d)", k), "Ancestry coefficient", file.path("outputs","allK_sNMF", sprintf("sNMF_K%d.png", k)), colpal_byK[[as.character(k)]]) }
```

---

# 5. Population-Level Mean Q (bestK)

```{r pop-mean-q}
Q_df <- as.data.frame(Q) %>% tibble::rownames_to_column("Individual") %>% dplyr::mutate(Population = as.character(popfac))
Q_means <- Q_df %>% dplyr::group_by(Population) %>% dplyr::summarise(dplyr::across(starts_with("K"), mean, na.rm = TRUE), .groups = "drop") %>% dplyr::arrange(Population)
readr::write_csv(Q_means, file.path("outputs", sprintf("%s_population_mean_Q_K%d.csv", params$save_prefix, bestK)))
Q_means
```

---

# 6. DAPC — fixed K (bestK) and **all K** versions to subfolder; optional auto-K by BIC

```{r dapc-block, eval=params$run_dapc}
library(adegenet)
set.seed(params$seed)

# Use X_qc (imputed & centered from filtered gl) to avoid NA/Inf and naming issues
ids_X <- rownames(X_qc)
ids_q <- rownames(Q_list[[as.character(bestK)]])
common_ids <- intersect(ids_X, ids_q)
if (length(common_ids) < length(ids_X)) {
  message(sprintf("Restricting DAPC to %d individuals shared with sNMF/bestK.", length(common_ids)))
}
X_common <- X_qc[match(common_ids, ids_X), , drop = FALSE]
Q_best_common <- Q_list[[as.character(bestK)]][match(common_ids, ids_q), , drop = FALSE]
# population factor for these individuals
popvec_common <- as.character(pop(gl))[match(common_ids, indNames(gl))]
popfac_common <- factor(popvec_common, levels = levels(popfac))
ord_common <- order(popfac_common)
X_common_ord <- X_common[ord_common, , drop = FALSE]
Q_best_ord2 <- Q_best_common[ord_common, , drop = FALSE]
pop_ord_common <- popfac_common[ord_common]

safe_dapc_plot <- function(Xmat, pop_ord_fac, Qref, k, out_png, title, joined, colpal) {
  # K-means on Xmat, then DAPC
  npc <- max(1, min(100, ncol(Xmat) - 1))
  fc <- find.clusters(Xmat, n.pca = npc, n.clust = k, choose.n.clust = FALSE)
  if (any(table(fc$grp) < 2)) { message(sprintf("Skipping DAPC K=%d due to singleton clusters.", k)); return(invisible(NULL)) }
  dap <- dapc(Xmat, grp = fc$grp, n.pca = npc, n.da = k - 1)
  opt <- optim.a.score(dap)
  dap <- dapc(Xmat, grp = fc$grp, n.pca = opt$best, n.da = k - 1)
  P <- dap$posterior
  if (!is.null(Qref) && ncol(Qref) == ncol(P)) {
    C <- suppressWarnings(cor(P, Qref, use = "pairwise.complete.obs"))
    perm <- integer(k); Cwork <- C
    for (step in seq_len(k)) { idx <- which(Cwork == max(Cwork, na.rm = TRUE), arr.ind = TRUE)[1, ]; i <- idx[1]; j <- idx[2]; perm[j] <- i; Cwork[i,] <- -Inf; Cwork[,j] <- -Inf }
    P <- P[, perm, drop = FALSE]; colnames(P) <- colnames(Qref)
  }
  plot_structure_like(P, pop_ord_fac, joined, title, "DAPC membership (posterior)", out_png, colpal)
  invisible(P)
}

# Colors for each K
colpal_byK <- lapply(K_range, function(k) if (requireNamespace("scales", quietly = TRUE)) scales::hue_pal()(k) else rainbow(k)); names(colpal_byK) <- as.character(K_range)

# bestK DAPC
safe_dapc_plot(X_common_ord, pop_ord_common, Q_best_ord2, bestK,
               file.path("outputs", sprintf("DAPC_compoplot_sorted_K%d.png", bestK)),
               sprintf("DAPC membership (Q-like; K = %d)", bestK), joined,
               colpal_byK[[as.character(bestK)]])

# ALL-K DAPC to subfolder
for (k in K_range) {
  Qk <- Q_list[[as.character(k)]]
  Qk_ord <- if (!is.null(Qk)) Qk[match(rownames(X_common_ord), rownames(Qk)), , drop = FALSE] else NULL
  try(safe_dapc_plot(X_common_ord, pop_ord_common, Qk_ord, k,
                     file.path("outputs","allK_DAPC", sprintf("DAPC_K%d.png", k)),
                     sprintf("DAPC membership (Q-like; K = %d)", k), joined,
                     colpal_byK[[as.character(k)]]), silent = TRUE)
}

# Optional: auto-K by BIC on X_common_ord
K_explore <- max(params$K_max, bestK + 5)
fc_auto <- find.clusters(X_common_ord, max.n.clust = K_explore, n.pca = max(1, min(100, ncol(X_common_ord) - 1)), choose.n.clust = FALSE)
K_auto <- nlevels(fc_auto$grp)
# BIC curve if available
bic_vals <- NULL; if (!is.null(fc_auto$Kstat)) bic_vals <- fc_auto$Kstat else if (!is.null(fc_auto$stat)) bic_vals <- fc_auto$stat
if (!is.null(bic_vals)) {
  bic_df <- data.frame(K = seq_along(bic_vals), BIC = as.numeric(bic_vals))
  p_bic <- ggplot(bic_df, aes(K,BIC)) + geom_line(color="#2c7fb8") + geom_point(size=2,color="#2c7fb8") + geom_vline(xintercept=K_auto, linetype=2,color="firebrick") + labs(title="DAPC model selection (BIC)", x="K", y="BIC") + theme_bw(base_size = 12)
  ggsave(file.path("outputs", "DAPC_BIC_explore.png"), p_bic, width = 7.5, height = 4, dpi = 150)
}
if (K_auto != bestK && all(table(fc_auto$grp) >= 2)) {
  safe_dapc_plot(X_common_ord, pop_ord_common, if (!is.null(Q_list[[as.character(K_auto)]])) Q_list[[as.character(K_auto)]][match(rownames(X_common_ord), rownames(Q_list[[as.character(K_auto)]])), , drop = FALSE] else NULL,
                 K_auto, file.path("outputs", sprintf("DAPC_compoplot_sorted_AUTO_K%d.png", K_auto)),
                 sprintf("DAPC membership (auto K by BIC; K = %d)", K_auto), joined,
                 if (!is.null(colpal_byK[[as.character(K_auto)]])) colpal_byK[[as.character(K_auto)]] else (if (requireNamespace("scales", quietly=TRUE)) scales::hue_pal()(K_auto) else rainbow(K_auto)))
}
```

---

# 7. Diagnostics & Tables

```{r diagnostics}
pop_table <- data.frame(Population = names(sort(table(pop(gl)), decreasing = TRUE)), N = as.integer(sort(table(pop(gl)), decreasing = TRUE)))
conf_summary <- data.frame(Threshold=c(0.7,0.8,0.9), Proportion=c(mean(assignments$MaxQ>=0.7), mean(assignments$MaxQ>=0.8), mean(assignments$MaxQ>=0.9)))
list(Individuals_per_Population = pop_table, Assignment_Confidence = conf_summary)
```

---

# 8. Reproducibility

```{r session-info}
cat(sprintf("Seed: %s\n", params$seed))
cat(sprintf("K range: %d..%d; repetitions: %d; alpha: %g\n", params$K_min, params$K_max, params$repetitions, params$alpha))
cat(sprintf("QC: missing_ind <= %.2f, missing_loc <= %.2f, MAF >= %.3f\n", params$missing_ind, params$missing_loc, params$maf_min))
sessionInfo()
```

---

# Appendix: How to Knit / Re-run Programmatically

```{r, eval=FALSE}
rmarkdown::render("PopulationStructure_sNMF_DAPC_AllK_fixNA2.Rmd",
  params = list(
    genind_rds = "genind.rds",
    missing_ind = 0.20,
    missing_loc = 0.10,
    maf_min = 0.01,
    K_min = 1,
    K_max = 12,
    repetitions = 10,
    alpha = 100,
    seed = 2025,
    save_prefix = "sNMF",
    run_dapc = TRUE
  ),
  clean = TRUE
)
```
