---
output: pdf_document
header-includes:
- \usepackage{pdfpages}
- \usepackage{lscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
--- 

<!-- Outputs as PDF, and allows for mixing landscape and portrain with header includes line -->
<!-- Only lines that should have to be modified by User.   -->
<!-- Point to baseline folder (with appropriate title to be carried over to report) and species. -->
<!-- Rmarkdown doesn't pull directly from environment, so easiest just to specify species here. -->


<!-- FOLLWING 3 LINES REQUIRE USER INTERVENTION. -->
<!-- folder is the unique folder (quoted) containing all the necessary files -->
<!-- species is a simple_pop_stats supported species (quoted) (eg. "chinook"; "coho"; "pink"; "eulachon"; "sockeye"; "chum") -->
<!-- datatype is either "SNP" or "microsat", same notation as used in simple_pop_stats environment. -->
<!-- min_loci is a numeric value [don't need to quote] that specifies what minimum number of loci the baseline was filtered on -->

```{r,echo=FALSE}
folder <- "2020_V1.0_2020-10-23_baseline"
species <- "coho"
datatype <- "SNP"
min_loci <- 150
```

<!-- Create apporpritate report title, and specify options to be carried through.   -->


```{r setup, include=FALSE}
suppressWarnings(library(readr))
getwd()
citations <- read_tsv(file="reference_files/citations.txt")
title_var <- paste(species,folder,sep="_")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = paste0("W:/9_PBT/01_",species,"/reference_databases/",folder,"/"))


str_break = function(x, width = 80L) {
  n = nchar(x)
  if (n <= width) return(x)
  n1 = seq(1L, n, by = width)
  n2 = seq(width, n, by = width)
  if (n %% width != 0) n2 = c(n2, n)
  substring(x, n1, n2)
}
```

<!-- Create report header.   -->

---
title: "`r title_var`"
author: "Molecular Genetics Lab - DFO Pacific"
date: "`r format(Sys.time(), "%Y-%m-%d")`"
---

<!-- Specify operator by drawing user name from the environment.   -->
This is a baseline benchmarking document. This document was prepared by:

```{r,echo=FALSE, comment=NA}
Sys.getenv("USERNAME")
```

## Baseline examined

<!-- Specify the baseline used.   -->
The data presented here is tied to the baseline:

```{r,echo=FALSE, comment=NA}
print(gsub("^.*/", "", getwd()))
```
<!-- Specify the species used.   -->
The species examined here is:
```{r,echo=FALSE, comment=NA}
print(species)
```
<!-- Specify the number of markers used. Option 3 in simple pop stats -->
The number and type of markers genotyped within the baseline is:
```{r,echo=FALSE, message=FALSE, comment=NA}
mark_num.fn <- list.files(paste0(getwd(),"/Baseline_summary/"),pattern="*number_of_markers.csv$",full.names = T,ignore.case=TRUE)
suppressWarnings(library(readr))
if (length(mark_num.fn)==1){
mark_num <- read_tsv(mark_num.fn)
print(mark_num$x)
} else {
  print("No marker number summary found")
}

```

<!-- The minimum number of successfully genotyped loci for an individual to be retained in the baseline  -->
The minimum number of successfully genotyped loci for an individual to be retained in the baseline:
```{r,echo=FALSE, message=FALSE, comment=NA}

print(min_loci)



```

<!-- Specify the highest tray number for samples in the baseline. Option 18 in simple pop stats -->
The highest tray number found in the baseline:
```{r,echo=FALSE, message=FALSE, comment=NA}
tray.fn <- list.files(paste0(getwd(),"/Baseline_summary/"),pattern="*highest_tray_number.txt$",full.names = T,ignore.case=TRUE)
suppressWarnings(library(readr))
if (length(tray.fn)==1){
tray <- read_tsv(tray.fn)
print(tray$Highest_Tray_Number)
} else {
  print("No Tray number summary found")
}
```



\newpage
## Citations
<!-- This will need to be a citation file that it pulls from based on species, TBD -->
If you use reports generated from this baseline, please cite:

```{r,echo=FALSE, comment=NA, message=FALSE}

suppressWarnings(library(dplyr))
tmp_species <- species
tmp_datatype <- datatype
cite <- filter(citations,species %in% tmp_species)
cite <- filter(cite,datatype %in% tmp_datatype)
if(length(cite>0)){
  for (i in 1:nrow(cite)){
    cat(paste0("Reference #",i,"\n"))
    cat(str_break(cite$citation[i]),sep="\n")
    cat("\n")
  }  
} else {
print("TBD")
}
```

\newpage
## MD5s of utilized files

<!-- Calculate MD5s for baseline files -->
The md5sum of the files used to generate the presented report are:

```{r,echo=FALSE, comment=NA}
lf <- list.files(getwd())
ld <- list.dirs(getwd(),recursive=F)
lf <- lf[match(setdiff(normalizePath(lf),normalizePath(ld)), normalizePath(lf))]
lf <- lf[!grepl(".pdf",lf)]
suppressWarnings(library(tools))
md5_lf <- as.data.frame(md5sum(lf),stringsAsFactors = FALSE)
md5_lf <- cbind(rownames(md5_lf), data.frame(md5_lf, row.names=NULL))
colnames(md5_lf) <- c("files","md5sum")
suppressWarnings(library(pander))
pander::pander(md5_lf,split.cell=c(55,35),split.table=Inf,missing="",use.hyphening=TRUE)

```
The following summaries were built off these files.

\newpage
## Summary file paths

<!-- Identify files used below -->
Results presented in this document are pulled from the following files:

Changes made between versions:
```{r,echo=FALSE, comment=NA}
fn.1 <- list.files(paste0(getwd(),"/Baseline_summary/"),pattern="*changes.txt$",full.names = T,ignore.case=TRUE)
if(length(fn.1>0)){
str_break(fn.1)
cat(str_break(fn.1), sep = '\n')
} else {
  print("No changes were detailed.")
}
```

Notes regarding specific collections:
```{r,echo=FALSE, comment=NA}
fn.2 <- list.files(paste0(getwd(),"/Baseline_summary/"),pattern="*notes.txt$",full.names = T,ignore.case=TRUE)
if(length(fn.2>0)){
str_break(fn.2)
cat(str_break(fn.2), sep = '\n')
} else {
  print("No notes were detailed.")
}
```

Baseline summary file:
```{r,echo=FALSE, comment=NA}
fn.3 <- list.files(paste0(getwd(),"/Baseline_summary/"),pattern="*.baseline_summary.txt$",full.names = T,ignore.case=TRUE)
if(length(fn.3>0)){
str_break(fn.3)
cat(str_break(fn.3), sep = '\n')
} else {
  print("No baseline summary was found.")
}
```


Dendrogram PDF. Please note - .Tre file is found in the same folder:
```{r,echo=FALSE, comment=NA}
fn.4 <- list.files(paste0(getwd(),"/Dendrogram/"),pattern="*.pdf$",full.names = T,ignore.case=TRUE)
if(length(fn.4>0)){
str_break(fn.4)
cat(str_break(fn.4), sep = '\n')
} else {
  print("No PDF-based dendrogram was found.")
}
```

Summary of Simulations. Please note - remaining simulations results are found in the same folder:
```{r,echo=FALSE, comment=NA}
fn.5 <- list.files(paste0(getwd(),"/100_sims/"),pattern="*100_stats_2.*.txt$",full.names = T,ignore.case=TRUE)
if(length(fn.5>0)){
str_break(fn.5)
cat(str_break(fn.5), sep = '\n')
} else {
  print("No 100% simulations file was found.")
}
```

Summary figure of Simulations:
```{r,echo=FALSE, comment=NA}
fn.6 <- list.files(paste0(getwd(),"/100_sims/"),pattern="*plot.pdf$",full.names = T,ignore.case=TRUE)
if(length(fn.6>0)){
str_break(fn.6)
cat(str_break(fn.6), sep = '\n')
} else {
  print("No 100% simulations summary figure was found.")
}
```



\newpage
## Version history

<!-- Detail any changes made between versions. Two column format, column 1 baseline name, column 2 notes; requires *changes.txt naming -->
Changes made between versions:

```{r, echo=FALSE, message=FALSE, comment=NA}
changes.fn <- list.files(paste0(getwd(),"/Baseline_summary/"),pattern="*changes.txt$",full.names = T,ignore.case=TRUE)
suppressWarnings(library(readr))
if (length(changes.fn)==1){
changes <- read_tsv(changes.fn)

suppressWarnings(library(pander))
pander::pander(changes,split.cell=c(20,60),split.table=Inf,missing="")
} else if (length(changes.fn)==0) {
  print("No changes noted with this baseline summary")
  
}
```

\newpage
##  Notes for specific collections:
<!-- Detail any Specific notes per collections. No set format yet, but requires *notes.txt naming, and collections need in a column named "collections" -->
```{r, echo=FALSE, message=FALSE, comment=NA}
notes.fn <- list.files(paste0(getwd(),"/Baseline_summary/"),pattern="*notes.txt$",full.names = T,ignore.case=TRUE)
suppressWarnings(library(readr))
if (length(notes.fn)==1){
notes <- read_tsv(notes.fn)
notes$collection <- gsub("_"," ",notes$collection)
suppressWarnings(library(pander))
pander::pander(notes,split.cell=c(20,55),missing="",split.table=Inf)
} else if (length(notes.fn)==0) {
  print("No additional notes with this baseline summary.")
  
}

```





\newpage
## Dendrogram
<!-- Dendrogram is exported from FigTree - will display exactly as is here, so make changes in FigTree -->

A summary dendrogram produced following simple_pop_stats recommended pathway and the genepop file listed above.

\includepdf[pages={-}]{"`r list.files(paste0(getwd(),"/Dendrogram/"),pattern="*.pdf$",full.names = T,ignore.case=TRUE)`"}


\newpage




## 100% Simulation summary

<!-- Files are produced by full_sim() and plot_summarize_100_sim()  -->
A summary of 100% simulation results produced using the full_sims() script found in simple_pop_stats.

\newpage
```{r, echo=FALSE,fig.cap = "Mean accuracy to collection (white), repunit (grey), or region (if applicable - dark grey)."}
suppressWarnings(library(knitr))
sim_sum.fn <- list.files(paste0(getwd(),"/100_sims/"),pattern="*.pdf$",full.names = T,ignore.case=TRUE)
include_graphics(sim_sum.fn)

```


\newpage

\newpage
\blandscape


\scriptsize
```{r, echo=FALSE, message=FALSE}
#setwd("W:/9_PBT/01_chinook/reference_databases/2020_V1.0_2020-08-05_baseline/")
sims.fn <- list.files(paste0(getwd(),"/100_sims/"),pattern="*100_stats_2.*.txt$",full.names = T,ignore.case=TRUE)
suppressWarnings(library(readr))
sims <- read_tsv(sims.fn)


if("collection" %in% colnames(sims)){
    sims$collection <- gsub("_"," ",sims$collection)
}

if("collection_scenario" %in% colnames(sims)){
    sims$collection_scenario <- gsub("_"," ",sims$collection_scenario)
}

names(sims) <- gsub("_"," ",names(sims))
sims <- sims[order(sims$repunit),]


suppressWarnings(library(pander))
suppressWarnings(library(koRpus))
pander::pander(sims,split.cell=18,split.table=Inf,missing="",use.hyphening=TRUE)

```
\normalsize

\elandscape


\newpage
## Baseline summary - individual collections, years and N
<!-- Files are produced by summarise_rubias_baseline()  -->
Run by using `summarise_rubias_baseline()`:
\footnotesize
```{r, echo=FALSE, message=FALSE}
#setwd("W:/9_PBT/01_chinook/reference_databases/2020_V1.0_2020-08-05_baseline/")
bsum.fn <- list.files(paste0(getwd(),"/Baseline_summary/"),pattern="*.baseline_summary.txt$",full.names = T,ignore.case=TRUE)
suppressWarnings(library(readr))
bsum <- read_tsv(bsum.fn)

bsum$Population <- gsub("_"," ",bsum$Population)


suppressWarnings(library(pander))
suppressWarnings(library(koRpus))
pander::pander(bsum,split.cell=18,split.table=Inf,missing="",use.hyphening=TRUE)

```
\normalsize

